---
title: "ISYE 6414: Homework 4 Peer Assessment Assignment"
author: ""
output:
  html_document:
    theme: cosmo
    highlight: haddock
    toc: true
    toc_depth: 3
---

```{r postprocess, include=F, echo=F, cache=F}
.path_r <- path.expand(rstudioapi::getSourceEditorContext()$path)
.path_sans_ext <- tools::file_path_sans_ext(.path_r)
.path_rmd <- paste0(.path_sans_ext, ".Rmd")
# rmarkdown::render(knitr::spin(.path_r, knit = FALSE))
# spelling::spell_check_files(.path_rmd)

```
```{r setup, include=F, cache=F}
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  # rows.print = 25,
  # rows.print = 25,
  echo = TRUE,
  # cache = FALSE,
  cache = TRUE,
  include = TRUE,
  fig.show = "asis",
  fig.align = "center",
  fig.width = 8,
  # size = "small",
  fig.height = 5,
  # fig.width = 5,
  # out.width = 5,
  # fig.asp = 0.75,
  warning = FALSE,
  message = FALSE
)
```

<style type="text/css">
body, td {
  font-size: 14px;
}
code.r{
  font-size: 12px;
}
pre {
  font-size: 12px;
}
instructions {
  font-style: italic;
}
response {
  font-weight: bold;
}
hide {
  display: none;
}
</style>

## 1. Circuit.

<instructions>
The system S consists of five independent elements $E_i$, $i = 1, \dots , E_5$, connected as in Figure 1. Probability that the element $E_i$ is operational at the end of time interval $[0,t]$ is given as
</instructions>

$$
p_i(t) = e^{-\lambda_i t}, t \geq 0,
$$
<instructions>
for $\lambda_1 = 1, \lambda_2 = 2, \lambda_3 = 1/2, \lambda_4 = 1/3, \lambda_5 = 1$.
</instructions>


<instructions>
(a) Find the probability that the system $S$ will be operational at time $t$. Plot this probability as a function of time $t$. What is this probability for $t = 1/2$?
</instructions>

<instructions>
(b) Find the probability that component $E_5$ was operational at time $t = 1/2$, if the system was operational at that time.
</instructions>

<instructions>
**Hint**: If you consider (b), it is conditional probability, more precisely, a posterior probability of the hypothesis $H_1 : E_5$ operational at time $t$, given that system S is operational at t. Thus, solve part (a) as a total probability with $H1$ and $H2 = H_1^C$ as hypotheses. Under the two hypotheses the system simplifies as in Figure 2 and it is easy to find $P(S|H1)$ and $P(S|H2)$.
Then (b) is just a Bayes formula.
</instructions>


### (a)

<response>
Using the notation of the hint, we have

$$
\begin{array}{lcl}
P(S_{H_1}|X) & = & 1 - P(E_1 \cup E_3) & = & 1 - (1 - P(E_1)) (1 - P(E_3)) \\
P(S_{H_2}|X) & = & & = & 1 - (1 - P(E_1) P(E_2)) (1 - P(E_3) P(E_4))
\end{array}
$$
</response>

### (b)

## 3. Classifier

<instructions>
In a machine learning classification procedure the items are classified as 1 or
0. Based on a training sample of size 120 in which there are 65 1’s and 55 0’s, the classifier
predicts 70 1’s and 50 0’s. Out of 70 items predicted by the classifier as 1, 52 are correctly
classified.
From the population of items where the proportion of 0-labels is 99% (and 1-labels 1%),
an item is selected at random. What is the probability that the item is of label 1, if the
classifier says it was.
</instruction>

Let $C$ represent correctness.
$P(C)=52/70$, $P(X=1)=65/120$, $P(X=0)=55/120$, $P(Y=1)=70/120$, $P(Y=0)=50/120$.


