---
title: 'ISYE 6420 Homework 3'
author: ''
output:
  html_document:
    theme: cosmo
    highlight: haddock
    toc: true
---

```{r setup, include=F, cache=F}
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  # rows.print = 25,
  # rows.print = 25,
  echo = TRUE,
  # cache = FALSE,
  cache = TRUE,
  include = TRUE,
  fig.show = "asis",
  fig.align = "center",
  fig.width = 8,
  # size = "small",
  fig.height = 5,
  # fig.width = 5,
  # out.width = 5,
  # fig.asp = 0.75,
  warning = FALSE,
  message = FALSE
)
```

```{r postprocess, include=F, echo=F, cache=F}
.path_sans_ext <- file.path('hws', 'hw02')
.path_rmd <- paste0(.path_sans_ext, '.Rmd')
# spelling::spell_check_files(.path_rmd)
```


<style type="text/css">
body, td {
  font-size: 16px;
}
code.r{
  font-size: 12px;
}
pre {
  font-size: 12px;
}
instructions {
  font-style: italic;
}
response {
  font-weight: bold;
}
hide {
  display: none;
}
.emphasized {
  font-size: 3em;
}
</style>

```{r setup-1, include=F, echo=F, eval=T}
library(tidyverse)
format_num <- function(x, digits = 4) {
  fmt <- sprintf("%%.%df", digits)
  sprintf(fmt, x)
}
```

# 1. Estimating the Precision Parameter of a Rayleigh Distribution.

## Instructions

<instructions>

</instructions>

## Response

# 2. Estimating Chemotherapy Response Rates

## Instructions

<instructions>
An oncologist believes that 90% of cancer patients will respond to a new chemotherapy treatment and that it is unlikely that this proportion will be below 80%. Elicit a beta prior on proportion that models the oncologist’s beliefs.<br/><br/>
Hint: For elicitation of the prior use $\mu = 0.9, \mu − 2 \sigma = 0.8$ and expressions for $\mu$ and $\sigma$. for beta.<br/><br/>
During the trial, in 30 patients treated, 22 responded.
(a) What are the likelihood and posterior distributions? What is the Bayes estimator of
the proportion?<br/>
(b) Using Octave, R, or Python, fine 95% Credible Set for $p$. <br/>
(c) Using Octave, R, or Python, test the hypothesis $H_0 : p \geq 2/3$ against the alternative $H_1 : p < 2/3$. <br/>
(d) Using WinBUGS, find the Bayes estimator, Credible Set and test, and compare results
with (a-c).
</instructions>

<hide>
(a)
likelihood -> 2018s-hw03-q2-a, ~2018s-hw03-q1-a
posterior ->  2018f-hw03-q1-iii, ~2018s-hw03-q1-a, ~2018s-hw03-q2-b
Bayes -> 2018f-hw03-q1-iii
(b) -> 2018f-hw03-q1-iii, 2018s-hw03-q2-c
(c) -> ~2018s-hw03-q1-c
(d) -> 2018f-hw03-q1-iii, ~2018s-hw03-q1-d, 2018s-hw03-q2-d
</hide>

## Response

### a

The expectation of a beta $\mathcal{Be}$ distribution 
for some random variable $X$ is $\operatorname{E}[X] = \frac{\alpha}{\alpha+\beta}$.
Given $\mu = 0.9, \mu − 2 \sigma = 0.8$, 
we can use algebra (or "trial and error")
to find that $\alpha = 0.72, \beta = 0.25 \alpha = 0.18$.

<hide>
the variance is $\operatorname{Var}(X) = \frac{\alpha \beta}{(\alpha + \beta)^2 (\alpha + \beta + 1)}$
</hide>

We can verify that these values are valid with `R`.

```{r hw03-q2-a-1, include=T, echo=T, eval=T}
alpha_0 <- 0.72
beta_0 <- 0.18
compute_beta_mu <- function(alpha, beta) {
  alpha / (alpha + beta)
}
mu_0 <- compute_beta_mu(alpha_0, beta_0)
mu_0
```

<hide>next, dwarf plants</hide>

Given the problem statement, 
<response>
we deduce the likelihood follows a binomial $\mathcal{Bin}$ distribution
We can model the likelihood distribution as follows.
</response>

$$
\begin{array}{c}
X \sim \mathcal{Bin}(n, k) \\
f_x(x|p) = \left( \begin{array}{c} n \\ k \end{array} \right) p^{x} (1 - p)^{n-k}
\end{array}
$$

The posterior distribution, $\pi (p | x)$,
is proportional to the product of the likelihood, $f_X(x|p)$, 
and the prior distribution, $\pi(p)$. Therefore, we have

$$
\begin{array}{lcl}
\pi(p | x) & \propto & f(x | p) \pi(p) \\
& = & p^{x}(1-p)^{n-x}
\end{array}
$$

<hide>next, continue dwarf plants</hide>

...

<response>
Thus, the posterior distribution looks as follows.
</response>

$$
\begin{array}{lcl}
p|x & \sim & \mathcal{Be}(\alpha', \beta') \\
& = & \mathcal{Be}(x + 1, n - x + 1)
\end{array}
$$

<response>
Given $n = 30$ patients and $x = 22$ who responded, we find that the posterior is
</response>

$$
\begin{array}{lcl}
\mathcal{Be}(x + 1, N - x +1) & = & \mathcal{Be}(20 + 1, 30 - 20 + 1) \\
& = &\mathcal{Be}(21, 11)
\end{array}
$$

We can verify that the posterior mean $\mu_1$ is less than the prior mean $\mu_0$.

```{r hw03-q2-a-2, include=T, echo=T, eval=T}
n <- 30
x <- 22
alpha_1 <- x - 1
beta_1 <- n - x + 1
mu_1 <- compute_beta_mu(alpha_1, beta_1)
mu_1
```

<response>
The Bayes estimator of the proportion is simply the expected value 
of the parameter under the posterior distribution
The expectation of the posterior is $\frac{21}{32} = 0.65625$.
</response>

$$
\begin{array}{lcl}
\mathcal{Be}(21, 11) & = & \frac{(21)}{(21)+(11)} \\
& = & \frac{21}{32} \\
& = & 0.65625.
\end{array}
$$

### b

The 95% equitailed credible set is based on the posterior distribution.
This means that the set $[q_{l}, q_{u}]$ contains the true
proportion of patient who respond with 95% probability.

In other words,

$$
\begin{array}{l}
\int_{-\infty}^{q_{l}} \pi(p|x) dp \leq \frac{\alpha}{2}, 
\int_{q_{u}}^{\infty} \pi(p|x) dp \leq \frac{\alpha}{2}
\text{ s.t. }
\Pr(p \in \left[q_{l}, q_{h}\right] | X) \geq 1 - \alpha
\end{array}
$$

where $\alpha = 0.95$ here.

<response>
We can calculate the credible set using `R` as shown below.
</response>

```{r hw03-q2-b-1, include=T, echo=T, eval=T}
compute_credible_set <- function(alpha, beta, level = 0.95) {
  stopifnot(level <= 1L)
  q_buffer <- (1 - level) / 2
  q_l <- (1 - level) - q_buffer
  q_u <- level + q_buffer
  res <-
    c(
      l = qbeta(q_l, shape1 = alpha, shape2 = beta),
      u = qbeta(q_u, shape1 = alpha, shape2 = beta)
    )
}

credible_set_1 <- compute_credible_set(alpha_1, beta_1)
credible_set_1
```

```{r hw03-q2-b-2, include=F, echo=F, eval=T}
credible_set_chr_1 <- format_num(credible_set_1)
credible_set_chr_1
```

We find that the 95% credible set is (`r credible_set_chr_1[1]`, `r credible_set_chr_1[2]`)

### c

```{r hw03-q2-c-1, include=F}
# 2018f-hw03-q2-a
pbeta(0.75, 137, 65) # `betacdf(3/4,137,65)` in Matlab
# 2018f-hw03-q2-b
qbeta(0.025, 137, 65) # `betainv(0.025,137,65)` in Matlab
qbeta(0.975, 137, 65) # `betainv(0.975,137,65)` in Matlab
```

To test the null hypothesis $H_0 : p \geq 2/3$ (against the alternative $H_1 : p < 2/3$),
we calculate the posterior probability of $p \geq 2/3$.

$$
\begin{array}{lcl}
\Pr(p < 2/3) & = & \int_{0}^{2/3} \mathcal{Be}(23, 11) dp \\
& = & \frac{\Gamma(23 + 11)}{\Gamma(23) \Gamma(11)} \int_{0}^{2/3} p^{23-1}(1-p)^{11-1} dp \\
& = & \frac{\Gamma(34)}{\Gamma(23) \Gamma(11)} \int_{0}^{2/3} p^{22}(1-p)^{10} dp \\
& \approx & 0.3297.
\end{array}
$$

We can verify this result with `R`.

```{r hw03-q2-c-1, include=T, echo=T, eval=T}
p_0 <- 2 / 3
p_h <- pbeta(p_0, alpha_1, beta_1)
p_h
```

```{r hw03-q2-c-2, include=F, echo=F, eval=T}
p_h_chr <- format_num(p_h, 4)
p_h_chr
alpha_h <- 1 - p_h
alph_h_chr <- format_num(alpha_h, 4)
alph_h_chr
```

Thus, at any significance level less than $\alpha \geq (1 - `r p_h_chr`) = `r alpha_h_chr`$,
we do not reject the null hypothesis(?).


