---
title: 'ISYE 6420: Homework 3'
author: ''
output:
  html_document:
    css: ../styles.css
    theme: cosmo
    highlight: haddock
    toc: true
---

```{r setup, include=F, cache=F}
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  # rows.print = 25,
  # rows.print = 25,
  echo = TRUE,
  # cache = FALSE,
  cache = TRUE,
  include = TRUE,
  fig.show = "asis",
  fig.align = "center",
  fig.width = 8,
  # size = "small",
  fig.height = 5,
  # fig.width = 5,
  # out.width = 5,
  # fig.asp = 0.75,
  warning = FALSE,
  message = FALSE
)
```

```{r postprocess, include=F, echo=F, cache=F}
.path_sans_ext <- file.path('hws', 'hw02')
.path_rmd <- paste0(.path_sans_ext, '.Rmd')
# spelling::spell_check_files(.path_rmd)
```

```{r setup-1, include=F, echo=F, eval=T}
library(tidyverse)
format_num <- function(x, digits = 4) {
  fmt <- sprintf("%%.%df", digits)
  sprintf(fmt, x)
}
```

# 1. Estimating the Precision Parameter of a Rayleigh Distribution.

## Instructions

<instructions>
If two random variables $X$ and $Y$ are independent of each other and 
normally distributed with variances equal to $\sigma^2$, then 
the variable $R = \sqrt{X^2+Y^2}$ follows the Rayleigh distribution.
Parameterized with precision parameter $\xi = \frac{1}{\sigma^2}$, 
the Rayleigh random variable $R$ has a density
</instructions>

$$
\begin{align}{c}
f(r) = \xi r \text{ exp } \{ \frac{-\xi r^2}{2} \}, \quad r \geq 0, \xi > 0.
\end{align}
$$

<instructions>
An example of such random variable would be the distance of darts from the target center
in a dart-throwing game where the deviations in the two dimensions of the target plane are
independent and normally distributed.<br/>
(a) Assume that the prior on $\xi$ is exponential with the rate parameter $\lambda$.
Show that the posterior is gamm $\mathcal{Ga}(2, \lambda+\frac{r^2}{2})$.<br/>
(b) Assume that $R_1 = 3, R_2 = 4, R_3 = 2, and R_4 = 5$ are Rayleigh-distributed
random observations representing the distance of a dart from the center. Find the posterior
in this case for the same prior form (a), and give a Bayesian estimate of $\xi$.
(c) Numerically find 95% Credible Set for $\xi$.
Hint: In (b) show that if $R_1, R_2, \dots, r_n$ are observed, and the prior on $\xi$ is exponential $\mathcal{E}(\lambda)$,
the the posterior is gamma $\mathcal{Ga}(n+1, \lambda+\frac{1}{2}\sum_{i=1}^{n} r_i^2)$.
</instructions>

## Response

# 2. Estimating Chemotherapy Response Rates

## Instructions

<instructions>
An oncologist believes that 90% of cancer patients will respond to a new chemotherapy treatment 
and that it is unlikely that this proportion will be below 80%. 
Elicit a beta prior on proportion that models the oncologist’s beliefs.<br/><br/>
Hint: For elicitation of the prior use $\mu = 0.9, \mu − 2 \sigma = 0.8$ 
and expressions for $\mu$ and $\sigma$. for beta.<br/><br/>
During the trial, in 30 patients treated, 22 responded.
(a) What are the likelihood and posterior distributions? What is the Bayes estimator of
the proportion?<br/>
(b) Using Octave, R, or Python, fine 95% Credible Set for $p$. <br/>
(c) Using Octave, R, or Python, test the hypothesis $H_0 : p \geq 2/3$ against the alternative $H_1 : p < 2/3$. <br/>
(d) Using WinBUGS, find the Bayes estimator, Credible Set and test, and compare results
with (a-c).
</instructions>

<hide>
(a)
likelihood -> 2018s-hw03-q2-a, ~2018s-hw03-q1-a
posterior ->  2018f-hw03-q1-iii, ~2018s-hw03-q1-a, ~2018s-hw03-q2-b
Bayes -> 2018f-hw03-q1-iii
(b) -> 2018f-hw03-q1-iii, 2018s-hw03-q2-c
(c) -> ~2018s-hw03-q1-c
(d) -> 2018f-hw03-q1-iii, ~2018s-hw03-q1-d, 2018s-hw03-q2-d
</hide>

## Response

### a

The expectation of a beta $\mathcal{Be}$ distribution 
for some random variable $X$ is $\operatorname{E}[X] = \frac{\alpha}{\alpha+\beta}$.
Given $\mu = 0.9, \mu − 2 \sigma = 0.8$, 
we can use algebra (or "trial and error")
to find that $\alpha = 0.72, \beta = 0.25 \alpha = 0.18$.

<hide>
the variance is $\operatorname{Var}(X) = \frac{\alpha \beta}{(\alpha + \beta)^2 (\alpha + \beta + 1)}$
</hide>

We can verify that these values are valid with `R`.

```{r hw03-q2-a-1, include=T, echo=T, eval=T}
alpha_0 <- 0.72
beta_0 <- 0.18
compute_beta_mu <- function(alpha, beta) {
  alpha / (alpha + beta)
}
mu_0 <- compute_beta_mu(alpha_0, beta_0)
mu_0
```

<hide>next, dwarf plants</hide>

Given the problem statement, 
<response>
we deduce the likelihood follows a binomial $\mathcal{Bin}$ distribution
We can model the likelihood distribution as follows.
</response>

$$
\begin{array}{c}
X \sim \mathcal{Bin}(n, k) \\
f_x(x|p) = \left( \begin{array}{c} n \\ k \end{array} \right) p^{x} (1 - p)^{n-k}
\end{array}
$$

The posterior distribution, $\pi (p | x)$,
is proportional to the product of the likelihood, $f_X(x|p)$, 
and the prior distribution, $\pi(p)$. Therefore, we have

$$
\begin{array}{lcl}
\pi(p | x) & \propto & f(x | p) \pi(p) \\
& = & p^{x}(1-p)^{n-x}
\end{array}
$$

<hide>next, continue dwarf plants</hide>

...

<response>
Thus, the posterior distribution looks as follows.
</response>

$$
\begin{array}{lcl}
p|x & \sim & \mathcal{Be}(\alpha', \beta') \\
& = & \mathcal{Be}(x + 1, n - x + 1)
\end{array}
$$

<response>
Given $n = 30$ patients and $x = 22$ who responded, we find that the posterior is
</response>

$$
\begin{array}{lcl}
\mathcal{Be}(x + 1, N - x +1) & = & \mathcal{Be}(22 + 1, 30 - 22 + 1) \\
& = &\mathcal{Be}(23, 9)
\end{array}
$$

<response>
The Bayes estimator of the proportion is simply the expected value 
of the parameter under the posterior distribution, which
is $\frac{23}{32} = 0.71875$.
</response>

$$
\begin{array}{lcl}
\mathcal{Be}(23, 9) & = & \frac{(23)}{(23)+(9)} \\
& = & \frac{23}{32} \\
& = & 0.71875.
\end{array}
$$

We can verify the value of the posterior mean $\mu_1$ using `R`.

```{r hw03-q2-a-2, include=T, echo=T, eval=T}
n <- 30
x <- 22
alpha_1 <- x + 1
beta_1 <- n - x + 1
mu_1 <- compute_beta_mu(alpha_1, beta_1)
mu_1
```

```{r hw03-q2-a-3, include=F, echo=F, eval=T}
mu_1_chr <- format_num(mu_1, 5)
mu_1_chr
```

### b

The 95% equitailed credible set is based on the posterior distribution.
This means that the set $[q_{l}, q_{u}]$ contains the true
proportion of patient who respond with 95% probability.

In other words,

$$
\begin{array}{l}
\int_{-\infty}^{q_{l}} \pi(p|x) dp \leq \frac{\alpha}{2}, 
\int_{q_{u}}^{\infty} \pi(p|x) dp \leq \frac{\alpha}{2}
\text{ s.t. }
\Pr(p \in \left[q_{l}, q_{h}\right] | X) \geq 1 - \alpha
\end{array}
$$

where $\alpha = 0.95$ here.

<response>
We can calculate the credible set using `R` as shown below.
</response>

```{r hw03-q2-b-1, include=T, echo=T, eval=T}
compute_credible_set <- function(alpha, beta, level = 0.95) {
  q_buffer <- (1 - level) / 2
  q_l <- (1 - level) - q_buffer
  q_u <- level + q_buffer
  res <-
    c(
      l = qbeta(q_l, shape1 = alpha, shape2 = beta),
      u = qbeta(q_u, shape1 = alpha, shape2 = beta)
    )
}

credible_set_1 <- compute_credible_set(alpha_1, beta_1)
credible_set_1
```

```{r hw03-q2-b-2, include=F, echo=F, eval=T}
credible_set_chr_1 <- format_num(credible_set_1)
credible_set_chr_1
```

We find that the 95% credible set is (`r credible_set_chr_1[1]`, `r credible_set_chr_1[2]`)

### c

```{r hw03-q2-c-0, include=F, echo=F, eval=F}
# 2018f-hw03-q2-a
pbeta(0.75, 137, 65) # `betacdf(3/4,137,65)` in Matlab
# 2018f-hw03-q2-b
qbeta(0.025, 137, 65) # `betainv(0.025,137,65)` in Matlab
qbeta(0.975, 137, 65) # `betainv(0.975,137,65)` in Matlab
```

To test the null hypothesis $H_0 : p \geq 2/3$ (against the alternative $H_1 : p < 2/3$),
we calculate the posterior probability of $p \geq 2/3$ using `R` as follows.

```{r hw03-q2-c-1, include=T, echo=T, eval=T}
p_0 <- 2 / 3
p_h <- pbeta(p_0, alpha_1, beta_1)
p_h
```

```{r hw03-q2-c-2, include=F, echo=F, eval=T}
p_h_chr <- format_num(p_h, 4)
p_h_chr
alpha_h <- 1 - p_h
alpha_h_chr <- format_num(alpha_h, 4)
alpha_h_chr
```

Analytically, the calculation looks as follows.

$$
\begin{array}{lcl}
\Pr(p < 2/3) & = & \int_{0}^{2/3} \mathcal{Be}(23, 9) dp \\
& = & \frac{\Gamma(23 + 9)}{\Gamma(23) \Gamma(9)} \int_{0}^{2/3} p^{23-1}(1-p)^{9-1} dp \\
& = & \frac{\Gamma(34)}{\Gamma(23) \Gamma(9)} \int_{0}^{2/3} p^{22}(1-p)^{10} dp \\
& \approx & `r alpha_h_chr`.
\end{array}
$$

Thus, at any significance level less than $\alpha \geq (1 - `r p_h_chr`) = `r alpha_h_chr`$,
we do not reject the null hypothesis(?).

### d

WinBUGS code to evaluate each of the following answers may look as follows.

```
model {
	x ~ dbin(p, n)
	p ~ dbeta(alpha, beta)
	h_1 <- step(2/3 - p)
}

# data
list(alpha = 0.72, beta = 0.18, x = 22, n = 30)
```

The output is 

<hide>
https://www.tablesgenerator.com/markdown_tables
</hide>

| node  | mean   | sd      | MC error | 2.5%   | median | 97.5%  | start | sample |
|-------|--------|---------|----------|--------|--------|--------|-------|--------|
| `h_1` | 0.1885 | 0.3911  | 0.001243 | 0.0    | 0.0    | 1.0    | 1     | 100000 |
| `p`   | 0.7352 | 0.07807 | 2.517E-4 | 0.5693 | 0.7404 | 0.8719 | 1     | 100000 |

The Bayes estimator corresponds to the mean of `p`, which is 0.7352.
This is very close to the value that we found analytically---`r mu_1`.

The 95% credible set corresponds to the 2.5% and 97.5% values for `p`, which
are 0.5693 and 0.8719 respectively.
These values are reasonably close to the values we found with `R`--`r credible_set_chr_1[1]` and `r credible_set_chr_1[2]`.

The probability of the alternative hypothesis $H_1$ corresponds to the
mean of the `h_1` variable---0.1885. This is relatively close to the
value that we found with `R`---`r p_h_chr`.


