---
title: 'ISYE 6420: Homework 2'
author: ''
output:
  html_document:
    styles: ../styles.css
    theme: cosmo
    highlight: haddock
    toc: true
---

```{r setup, include=F, cache=F}
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  # rows.print = 25,
  # rows.print = 25,
  echo = TRUE,
  # cache = FALSE,
  cache = TRUE,
  include = TRUE,
  fig.show = "asis",
  fig.align = "center",
  fig.width = 8,
  # size = "small",
  fig.height = 5,
  # fig.width = 5,
  # out.width = 5,
  # fig.asp = 0.75,
  warning = FALSE,
  message = FALSE
)
```

```{r postprocess, include=F, echo=F, cache=F}
.path_sans_ext <- file.path('hws', 'hw02')
.path_rmd <- paste0(.path_sans_ext, '.Rmd')
# spelling::spell_check_files(.path_rmd)
```

```{r setup-1, include=F, echo=F, eval=T}
library(tidyverse)
format_num <- function(x, digits = 4) {
  fmt <- sprintf("%%.%df", digits)
  sprintf(fmt, x)
}
```

# 1. Cell Clusters in 3D Petri Dishes

## Instructions

<instructions>
The number of cell clusters in a 3D Petri dish has a Poisson distribution with
mean $\lambda = 5$. Find the percentage of Petri dishes that have (a) 0
clusters, (b) at least one cluster, (c) more than 8 clusters, and (d) between 4
and 6 clusters inclusive.
</instructions>

## Response

Let $X$ represent the number of cell clusters. The instructions indicate that
$X \sim \operatorname{Pois}(\lambda=5)$.
Thus the probability mass function (pmf) $f(x)$ of $X$ is as follows.

$$
f(x) = \frac{\lambda{x} e^{-\lambda}}{x!} = \frac{5^{x} e^{-5}}{x!}.
$$

Note that $f(x) = P(X = x)$.

The cumulative distribution function (cdf) $F(x)$ of $X$ is as follows.

$$
F(x) = e^{-\lambda} \sum_{k=0}^{x} \frac{\lambda^k}{k!} = e^{-5} \sum_{k=0}^{x} \frac{5^k}{k!}.
$$

Note that $F(x) = P(X \leq x)$.

Now we address each part of the instructions.

### (a)

We use the pdf to find the percentage of Petri dishes having 0 clusters.

$$
P(X = 0)=\frac{5^{0} e^{-5}}{0!} = e^{-5} = 0.0067
$$

We can check this number using `R` as follows.

```{r hw02-q1-a-1, include=T, echo=T, eval=T}
l <- 5
p_q1a <- ppois(0, lambda = l, lower.tail = T)
p_q1a
```

<response>
Thus, we find that the percentage is 100 * `r format_num(p_q1a, 4)` = `r format_num(100 * p_q1a, 4)`.
</response>


### (b)

Next, we deduce the percentage of Petri dishes having at least one cluster from
our finding in (a).

$$
P(X \geq 1)= 1 - P(X = 0) = 1 - (e^{-5}) = 0.9933
$$

Again, we can check our number with `R`.

```{r hw02-q1-b-1, include=T, echo=T, eval=T}
1 - p_q1a
# Or
p_q1b <- ppois(0, lambda = l, lower.tail = F)
p_q1b
```

<response>
Thus, we find that the percentage is `r format_num(100 * p_q1b, 4)`%.
</response>

### (c)

To find the percentage of Petri dishes having more than 8 cluster, we use the cdf.

$$
P(X > 8) = 1 - P(X \leq 8) = 1 - (e^{-5} \sum_{k=0}^{8} \frac{5^{k}}{k!}).
$$

Let's use `R` to calculate this (since writing out all of the individual terms
in the sum would be tedious).

```{r hw02-q1-c-1, include=T, echo=T, eval=T}
p_q1c <- ppois(8, lambda = l, lower.tail = F)
p_q1c
```

<response>
We find that the percentage is `r format_num(100 * p_q1c, 4)`%.
</response>

### (d)

Finally, to find the percentage of Petri dishes having between 4 and 6 clusters
inclusive, we use the difference in cdfs. (Note that we must be careful with the
lower bound 4, which we want to include.)

$$
P(4 \leq X \leq 6) = P(X \leq 6) - P(X < 3) = (e^{-5} \sum_{k=0}^{6} \frac{5^{k}}{k!}) - (e^{-5} \sum_{k=0}^{3} \frac{5^{k}}{k!}) = e^{-5} \sum_{k=4}^{6} \frac{5^{k}}{k!}.
$$

```{r hw02-q1-d-1, include=T, echo=T, eval=T}
p_q1d <- ppois(6, lambda = l, lower.tail = T) - ppois(4 - 1, lambda = l, lower.tail = T)
p_q1d
```

<response>
We find that the percentage is `r format_num(100 * p_q1d, 4)`%.
</response>

# 2. Silver-Coated Nylon Fiber

## Instructions

<instructions>
Silver-coated nylon fiber is used in hospitals for its anti-static electricity
properties, as well as for antibacterial and antimycotic effects. In the
production of silver-coated nylon fibers, the extrusion process is interrupted
from time to time by blockages occurring in the extrusion dyes. The time in
hours between blockages, $T$, has an exponential $\mathcal{E}$(1/10) distribution,
where 1/10 is the rate parameter. <br/>
Find the probabilities that </br>
(a) a run continues for at least 10 hours,<br/>
(b) a run lasts less than 15 hours, and<br/>
(c) a run continues for at least 20 hours, given that it has lasted 10 hours.<br/><br/>
If you use software, be careful about the parametrization of exponentials.
</instructions>

## Response

The instructions indicate that $T \sim \mathcal{E}(1/\lambda=1/10)$.
(Note that $\lambda$ represents a rate. It is like the inverse of 
$\lambda$ for the Poisson distribution.)
The probability density function (pdf) and cdf are as follows. (Note that pdf
is appropriate for continuous distrubitons, while pmf is appropriate for discrete
distrubitions.

$$
f(x) = \lambda e^{-\lambda x} = - \frac{1}{10} e^{-x/10}, x \geq 0.
$$

$$
F(x) = 1 - e^{-\lambda x} = 1 - e^{-x/10}, x \geq 0.
$$

### (a)

We find the probability of a run continuing for at least 10 hours as follows.

$$
P(X > 10) = 1 - P(X \leq 10) = 1 - (1 - e^{-(10)/(10)}) = e^{-1} = 0.3679.
$$

We can check this with `R`.

```{r hw02-q2-a-1, include=T, echo=T, eval=T}
exp(-1)
# Or
l <- 10
r <- 1 / l
p_q2a <- pexp(10, rate = r, lower.tail = F)
p_q2a
```

<response>
The probability is `r format_num(p_q2a, 4)`.
</response>

### (b)

We find the probability that a run lasts less than 15 hours as follows.

```{r hw02-q2-b-1, include=T, echo=T, eval=T}
1 - pexp(15, rate = r, lower.tail = F)
# Or
p_q2b <- pexp(15, rate = r, lower.tail = T)
p_q2b
```

<response>
The probability is `r format_num(p_q2b, 4)`.
</response>


### (c)

Note that the $\mathcal{E}$ distribution has the memoryless property, which means that 
$P(X > (s + t) | X > s) = P(X > t)$ for $s, t > 0$.


Now we calculate the probability that run continues for at least 20 hours, given that it has lasted 10 hours, as follows.

$$
P(X > 20 | X > 10) = P(X > 10) = e^{-10/10} = e^{-1} = 0.3679.
$$

Checking with `R`, we find the following.

```{r hw02-q2-c-1, include=T, echo=T, eval=T}
exp(-1)
# Or 
p_q2c <- pexp(20 - 10, rate = r, lower.tail = F)
p_q2c
```

<response>
The probability is `r format_num(p_q2c, 4)`.
</response>

# 3. 2-D Density Tasks

<instructions>
If
</instructions>
<hide>
$$
f(x,y) =
\begin{cases}
\lambda^2 e^{-\lambda y} & \quad 0 \leq x \leq y, & \lambda > 0 \\
0, & \quad \text{else}.
\end{cases}
$$
</hide>


$$
f(x,y) =
\left\{
\begin{array}{cll}
\lambda^2 e^{-\lambda y} & \quad 0 \leq x \leq y, & \lambda > 0 \\
0, & \quad \text{else}.
\end{array}
\right.
$$
<instructions>
Show that:<br/>
(a) marginal distribution $f_X(x)$ is exponential $\mathcal{E}(\lambda)$.<br/>
(b) marginal distribution $f_Y(y)$ is Gamma $\mathcal{G}a(2,\lambda)$.<br/>
(c) conditional distribution $f(y|x)$ is shifted exponential $f(y|x) = \lambda e^{-lambda(y-x)}$, $y \geq x$.
(d) condition distribution $f(x|y)$ is uniform $\mathcal{U}(0,y)$.
<instructions/>

## (a)


```{r hw02-q3-a-1, include=F, echo=F, eval=F}
l <- 1
f <- expression(l^2 * exp(-l * x))
# f <- expression(exp(-x))
d <- stats::D(f, 'x')
d

integrand <- function(x) eval(f)
lower <- 0
upper <- 1
res <- integrate(f = integrand, lower = lower, upper = upper)
res
```

<hide>
https://www.symbolab.com/solver/integral-calculator/%20%5Cint_%7Bx%7D%5E%7B%5Cinfty%7D%20C%5E%7B2%7De%5E%7B-Cy%7Ddy
</hide>

$$
\begin{array}{lcll}
f_X(x) & = & \int_{x}^{\infty} \lambda^2 e^{-\lambda y} dy \\
& = & \lambda^2 \int_{x}^{\infty} e^{-\lambda y} dy \\
& = & \lambda^2 (\frac{1}{-\lambda} e^{-\lambda y} \Big|_{x}^{\infty}) \\
& = & -\lambda (0 - e^{-\lambda x}) \\
& = & \lambda e^{-\lambda x}
\end{array}
$$


Note that the pdf of $\mathcal{E}(\lambda)$ is $f(x) = \lambda e^{-\lambda x}$,
which matches the result above.

## (b)

<hide>
https://www.symbolab.com/solver/integral-calculator/%20%5Cint_%7B0%7D%5E%7By%7D%20C%5E%7B2%7De%5E%7B-Cy%7Ddx
</hide>

$$
\begin{array}{lcl}
f_Y(y) & = & \int_{x0}^{y} \lambda^2 e^{-\lambda y} dx \\
& = & \lambda^2 e^{-\lambda y} \int_{0}^{y} dx \\
& = & \lambda^2 e^{-\lambda y} x \Big|_{0}^{y} \\
& = &\lambda^2 e^{-\lambda y} y
\end{array}
$$

Note that general form of the pdf of $\mathcal{G}a(\alpha,\beta)$ is as follows.

$$
f(x;\alpha,\beta) = \frac{\beta^{\alpha }x^{\alpha -1}e^{-\beta x}}{\Gamma(\alpha )},\quad x, \alpha ,\beta >0
$$

Then, if we substitute $\lambda$ for $\beta$ and 2 for $\alpha$, and
we evaluate the denominator as $\Gamma(2) = (2 - 1)! = 1$
(given the fact that $\Gamma(n) = (n - 1)! \quad  \forall \mathbb{Z}^+$),
then we see how the formulation above matches the distribution
that we should see--$\mathcal{G}a(2,\lambda)$.

$$
\begin{array}{lcl}
f(y) & = & \frac {\lambda^{2}y^{2-1}e^{-\lambda y}}{\Gamma(2)} \\
& = & \frac {\lambda^{2} y e^{-\lambda y}}{1} \\
& = & \lambda^{2} y e^{-\lambda y} \sim \mathcal{G}a(2,\lambda)
\end{array}
$$

## (c)

$$
\begin{array}{lcl}
f_{Y|X}(y|x) & = & \frac{f_{X,Y}(x,y)}{f_X(x)} \\
& = & \frac{\lambda^2 e^{-\lambda y}}{-\lambda (e^{-\lambda y} - 1)} \\
\end{array}
$$

(Note that we are being "verbose" by writing $X$ and $Y$ in the subscript of terms
that involve both, e.g. $f_{X,Y}(x,y)$ instead of just $f(x,y)$.

## (d)

$$
\begin{array}{lcl}
f_{X|Y}(x|y) & = & \frac{f_{X,Y}(x,y)}{f_Y(y)} \\
& = &  \\
\end{array}
$$

# 4. Nylon Fiber Continued.

<instructions>
In the Exercise 2, the times (in hours) between blockages of the 
extrusion process, $T$, had an exponential $\mathcal{E}(\lambda)$ distribution. 
Suppose that the rate parameter $\lambda$ is unknown, but there are three measurements of interblockage times, $T_1 = 3, T_2 = 13$, and $T_3 = 8$.<br/><br/>
(a) How would classical statistician estimate $\lambda$?<br/>
(b) What is the Bayes estimator of $\lambda$ if the prior is $\pi(\lambda) = \frac{1}{\sqrt{\lambda}}$, $\lambda > 0$? <br/><br/>
Hint: In (b) the prior is not a proper distribution, but the posterior is. Identify the posterior from the product of the likelihood from (a) and the prior, no need to integrate.
<instructions/>

## a.

A classical statistician would estimate $\lambda$ using the MLE, which is $\hat{\lambda}_{MLE} = \frac{1}{\overline{X}}$.
$$
\overline{X} = \frac{1}{n} \sum_{i=1}^n T_i = \frac{1}{3} (3 + 13 + 8) = 8.
$$

Thus,

$$
\hat{\lambda}_{MLE} = \frac{1}{\overline{X}} = \frac{1}{8}.
$$

## b.

The instruction indicate that we are trying to estimate the rate $T_1, \dots, T_n \sim \mathcal{E}(\lambda)$ given the prior $\pi(\lambda)$. (Here, $n = 3$.) 
Moreover, note that Bayes formula can be written generally
(in terms of the pdf $f$. the marginal $m$, and the prior $pi(\lambda)$
as follows (using our notation of $T_i$, which takes on values of $t_i$ and $\lambda$).

$$
\pi(\lambda|t_1,\dots,t_n) = \frac{f(t_1,...,t_n|\lambda) \pi(\lambda)}{m(t_1,\dots,t_n)}.
$$

<hide>
**NOTE TO SELF**: This is not right. (This question is not really analogous to the 2018F question.)
</hide>


From the literature, we know that the gamma distribution $\operatorname{Gamma}(\alpha,\beta)$
is the conjugate prior for the exponential distribution. Further, we can
write out the posterior distribution with respect to
the prior distribution as follows.

$$
\pi(\lambda|t_i,\dots,t_n) = \operatorname{Gamma}(\pi(\lambda),\alpha,\beta)
$$

The parameter $\alpha$ represents the number of prior observations, and the
parameter $\beta$ represents the sum of the prior observations.
</hide>
