---
title: 'ISYE 6420: Homework 4'
author: 'aelhabr3'
output:
  html_document:
    css: ../styles_hw.css
    theme: cosmo
    highlight: haddock
    toc: false
---

```{r setup, include=F, cache=F}
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  # rows.print = 25,
  # rows.print = 25,
  echo = TRUE,
  # cache = FALSE,
  cache = TRUE,
  include = TRUE,
  fig.show = 'asis',
  fig.align = 'center',
  fig.width = 8,
  # size = "small",
  fig.height = 5,
  # fig.width = 5,
  # out.width = 5,
  # fig.asp = 0.75,
  warning = FALSE,
  message = FALSE
)
```

```{r postprocess, include=F, echo=F, cache=F}
.path_sans_ext <- file.path('hw4')
.path_rmd <- paste0(.path_sans_ext, '.Rmd')
# spelling::spell_check_files(.path_rmd)
```

```{r setup-1, include=F, echo=F, eval=T}
library(tidyverse)
format_num <- function(x, digits = 4) {
  fmt <- sprintf('%%.%df', digits)
  sprintf(fmt, x)
}
```

```{r pdf-convert-1, include=F, echo=F, eval=F}
convert_page_to_lines <- function(page) {
  page %>%
    str_split('\\n') %>%
    purrr::map(str_squish) %>%
    unlist() %>%
    enframe(name = 'idx_line', value = 'line')
}

unnest_pages <- function(pages) {
  pages %>%
    mutate(lines = purrr::map(page, convert_page_to_lines)) %>%
    select(-page) %>%
    unnest(lines)
}
path_hw <- 'Homework4f.pdf'
pages <- 
  path_hw %>% 
  tibble(path = .) %>% 
  mutate(
    page = purrr::map(path, ~pdftools::pdf_text(.))
  ) %>%
  select(-path)
lines <- pages %>% unnest_pages()
lines
lines_text <-
  lines %>%
  pull(line) %>% 
  paste(collapse = ' ', sep = '')
lines_text
```

# 1. Metropolis: The Bounded Normal Mean.

## Instructions

<instructions>
Suppose that we have information that the normal mean $\theta$ is bounded between $−m$ and $m$, for some known number $m$. In this case it is natural to elicit a prior on $\theta$ with the support on interval $[−m, m]$. A prior with interesting theoretical properties supported on $[−m, m$] is Bickel-Levit prior, 
</instructions>

$$
\begin{array}{c}
\pi(\theta)=\frac{1}{m} \cos ^{2}\left( \frac{\pi \theta}{2 m}\right), \quad -m \leq \theta \leq m.
\end{array}
$$

<instructions>
Assume that a sample $[−2, −3, 4, −7, 0, 4]$ is observed from normal distribution 
</instructions>

<instructions>
with a known precision $\tau = 1/4$. Assume also that the prior on $\theta$ is Bickel-Levit, with $m = 2$. This combination likelihood/prior does not result in an explicit posterior (in terms of elementary functions). Construct a Metropolis algorithm that will sample from the posterior of $\theta$.<br/><br/>
(a) Simulate 10,000 observations from the posterior, after discarding first 500 observations (burn-in), and plot the histogram of the posterior.<br/>
(b) Find Bayes estimator of $\theta$, and 95% equitailed Credible Set based on the simulated observations.<br/><br/>
Suggestions:<br/>
(i) Take uniform distribution on $[−m, m]$ as a proposal distribution since it is easy to sample from. This is an independence proposal, the proposed $\theta'$ does not depend on the current value of the chain, $\theta$.<br/>
(ii) You will need to calculate $\sum_{i=1}^n (y_i − \theta)^2$ for current $\theta$ and $\sum_{i=1}^n i=1 (y_i − \theta')^2$ for the proposed $\theta'$, prior to calculating the Metropolis ratio.
</instructions>


$$
\begin{array}{c}
f(y|\theta) \propto \sqrt{\tau} \text{ exp } \{ \frac{\tau}{2} (y - \theta) ^2 \},
\end{array}
$$

## Response

### a

As noted in the instructions, the posterior cannot be written easiliy in terms of "elementary" functions.
Alternatively, the Metropolis algorithm can be used to approximate a posterior distribution.

First, let's construct the target density distribution. This is proportional to the posterior, which is the product of the prior and the likelihood. Using the Bickel-Levit prior $\pi(\theta)$ with $m = 2$ and the likelihood distribution $f(y|\theta)$ with $\tau = \frac{1}{4}$, <hide>and $\bar{y} = \frac{1}{n} \sum_{i=1}^n} y_i = \frac{1}{6} (-6) = -1$</hide>, we formulate the target density function as follows.

$$
\begin{array}{rcl}
\text{ posterior } & \propto & \text{ prior } \times \text{ likelihood } \\
\pi(y|\theta) & \propto & \pi(\theta) \times f(y|\theta) \\
& = & (\frac{1}{(2)} \cos ^{2}\left( \frac{\pi \theta}{2 (2)}\right)) (\sqrt{(1/4)} \text{ exp } \{ \frac{(1/4)}{2} (y - \theta) ^2 \}) \\
\end{array}
$$

(This is still a relatively complex expression, but code can help us solve for given values of $\theta$ and $\tau$.)

Using the suggestion and given $m=2$, we define the proposal density as $\mathcal{U}(-2, 2)$.

Then we have the acceptance probability $\rho$ as


$$
\begin{array}{c}
\rho\left(\theta_{n}, \theta^{\prime}\right)=
\min \left\{1, \frac{\pi\left(\theta^{\prime}\right)}{\pi\left(\theta_{n}\right)} \frac{q\left(\theta_{n} | \theta^{\prime}\right)}{q\left(\theta^{\prime} | \theta_{n}\right)}\right\}
=\min \left\{1, z \right\}.
\end{array}
$$

Note that we implement the Metropolis algorithm in the code as follows.

1. Start with arbitary $\theta$.[^1]

2. At stage $n$, generate proposal $\theta^{\prime}$ from $q(y|x_n)$.

3. Define

$$
\theta_{n+1} = \begin{cases}
\theta^{\prime} & \text{ with probability } & \rho(\theta_{n}, \theta^{\prime}), \\
\theta_{n} & \text{ with probability } &  1 - \rho(\theta_{n}, \theta^{\prime}).
\end{cases}
$$

(Generate $U = \mathcal{U}(0, 1)$ and accept proposal $\theta^{\prime}$ if $U \leq \rho(\theta_{n}, \theta^{\prime})$.)

4. Increase $n$ and go to Step 2.

[^1]: We don't really have an idea of the form of the support of the target; otherwise, we would try to initialize $\theta$ to be like the target.

See the code below for the implementation. (Note that `theta_prop` below represents $\theta^{\prime}$ in the formulations above.)

```{r q1-a-0, include=F, echo=F, eval=F}
# f_likelihood_simple <- function(theta) {
#   exp(theta^2 / 8)
# }
# dbl <- function(theta, m) {
#   theta <- punif(theta, min = -m, max = m)
#   bl(theta, m)
# }
# 
# rbl <- function(n, m) {
#   theta <- runif(n = n, min = -m, max = m)
#   bl(theta, m)
# }

# idx_final <- (n_burnin + 1):n_mcmc
# # Or?
# n_mcmc <- n_mcmc + n_burnin
# idx_final <- (n_burnin + 1):n_mcmc

# Data.
# ~~These are used to calculate the posterior likelihood, but not for MCMC.~~
# y <- ''

```

```{r q1-a-1, include=T, echo=T, eval=T}
library(tidyverse)

# Data, constants, hyperparameters, and helper functions.
y <- c(-2, -3, 4, -7, 0, 4)

n_mcmc_q1 <- 10000L
n_burnin_q1 <- 500L

m <- 2
tau <- 1 / 4


.f_likelihood_1 <- function(y, theta, tau) {
  sqrt(tau) * exp(-0.5 * tau * (y - theta)^2)
}

.f_likelihood_2 <- function(y, theta, tau) {
  sqrt(tau) * exp(-0.5 * tau * sum(y - theta)^2)
}

bl <- function(theta, m) {
  stopifnot(theta <= m || theta >= -m)
  (1 / m) * (cos((pi * theta) / (2 * m))) ^ 2
}

f_prior <- function(theta, .m = m) {
  bl(theta, m = .m)
}

f_proposal <- function(.n = 1, .m = m) { 
  runif(n = .n, min = -.m, max = .m) 
}

# f_likelihood_1 <- function(theta, .y = y, .tau = tau) {
#   sum(.f_likelihood_1(theta, y = y, tau = tau))
# }

f_likelihood_2 <- function(theta, .y = y, .tau = tau) {
  .f_likelihood_2(theta, y = y, tau = tau)
}

f_likelihood <- f_likelihood_2

is_likeinteger <- function(x, tol = .Machine$double.eps^0.5) {
  abs(x - round(x)) < tol
}

```
```{r q1-a-2-debug-1, include=F, echo=F, eval=T}
.x <- seq(-2, 2, by = 0.01)
qsave <- function(x, nm = deparse(substitute(x))) {
  ggplot2::ggsave(
    filename = sprintf('%s.png', nm),
    plot = x,
    units = 'in', 
    width = 7, 
    height = 5
  )
}
viz_prior <- tibble(x = .x, prior = f_prior(.x)) %>% qplot(x, prior, data = .)
viz_prior
qsave(viz_prior)
viz_prop <- tibble(x = .x, prop = f_proposal(.x)) %>% qplot(x, prop, data = .)
viz_prop
qsave(viz_prop)
viz_likl <- tibble(x = .x, likl = purrr::map_dbl(.x, f_likelihood)) %>% qplot(x, likl, data = .)
viz_likl
qsave(viz_likl)
tibble(x = .x, likl = purrr::map_dbl(.x, ~.f_likelihood_2(.x, y = -3, tau = tau))) %>% 
  qplot(x, likl, data = .)

```


```{r q1-a-2-debug-2, include=F, echo=F, eval=T}
.rnorm_int <- function(...) {
  as.integer(rnorm(...))
}

n_obs <- length(y)

rnorm_int <- function(n = n_obs, mean = 0, sd = 4) {
  # sort(.rnorm_int(n = n, mean = mean, sd = sd))
  .rnorm_int(n = n, mean = mean, sd = sd)
}

.seq <- c(1:9 / 10, 1:2 / 1) %>% sort()
# .y_1 <- rnorm_int()
.y_1 <- c(-3, -4, 0, 1, 7, 0) # Just a manual emulation of `.y`.
set.seed(1)
.y_2 <- rnorm_int()
mean(.y_2)
sd(.y_2)
# .y_2 <- c(3, -4, 1, -6, 1, 5)
res_sim <-
  c(-.seq, 0, .seq) %>% 
  sort() %>% 
  tibble(theta = .) %>% 
  mutate(idx = row_number()) %>% 
  select(idx, everything()) %>% 
  group_by(idx) %>% 
  mutate(
    # pi_0 = bl(theta, m = 4),
    pi = f_prior(theta),
    # f_1 = sum(.f_1(theta)), 
    # f = f_likelihood(theta),
    # f = purrr::map_dbl(theta, .f_likelihood_2, y = y, tau = tau),
    f_0 = .f_likelihood_2(theta, y = y, tau = tau),
    # f_0 = .f_likelihood_2(theta, y = sort(y), tau = tau),
    f_1 = .f_likelihood_2(theta, y = .y_1, tau = tau),
    f_2 = .f_likelihood_2(theta, y = .y_2, tau = tau)
  ) %>% 
  ungroup()
res_sim
res_sim %>% 
  gather(key = 'key', value = 'value', -matches('idx|theta')) %>% 
  ggplot() +
  aes(x = theta, y = value, color = key) +
  geom_point() +
  geom_line() +
  # scale_x_continuous(trans = 'reciprocal')
  xlim(-1.1, 1.1) +
  theme_light()
```

```{r q1-a-4, include=T, echo=T, eval=T}
# Main function.
.do_mcmc_mh_q1 <- function(..., n_mcmc = 10000L, n_burnin = 1000L, theta_init = 2) {
  
  # n_mcmc = n_mcmc_q1; n_burnin = n_burnin_q1; theta_init = 0
  
  stopifnot(is_likeinteger(n_mcmc))
  if(!is.null(n_burnin)) {
    stopifnot(is_likeinteger(n_burnin))
    stopifnot(n_burnin < n_mcmc)
  }
  cols_mat_mcmc <-
    c(
      'theta_current', 
      'theta_prop',
      'q_current',
      'q_prop',
      'pi_current',
      'pi_prop',
      'ratio_num',
      'ratio_den',
      'ratio', 
      'rho',
      'u', 
      'is_accepted'
    )
  mat_mcmc <- matrix(nrow = n_mcmc, ncol = length(cols_mat_mcmc))
  colnames(mat_mcmc) <- cols_mat_mcmc
  
  # Step 1 achieved with `.theta_init`.
  theta_current <- theta_init
  for (i in 1:n_mcmc) {
    #  i <- 1
    
    # Step 2.
    # set.seed(42)
    theta_prop <- f_proposal()
    
    # Step 3.
    q_current <- f_prior(theta = theta_current)
    q_prop <- f_prior(theta = theta_prop)
    
    pi_current <- f_likelihood_2(theta = theta_current)
    pi_prop <- f_likelihood_2(theta = theta_prop)
    
    ratio_num <- pi_prop * q_current
    ratio_den <- pi_current * q_prop
    ratio <- ratio_num / ratio_den
    
    u <- runif(n = 1, min = 0, max = 1)
    rho <- min(1, ratio)
    
    # Step 4.
    is_accepted <- 0
    if(u <= rho) {
      theta_current <- theta_prop
      is_accepted <- 1
    } else {
      # NULL
    }
    mat_mcmc[i, ] <- 
      c(
        theta_current,
        theta_prop,
        q_current,
        q_prop,
        pi_current,
        pi_prop,
        ratio_num,
        ratio_den,
        ratio,
        rho, 
        u, 
        is_accepted
      )
  }
  res_mcmc <- 
    mat_mcmc %>% 
    as_tibble() %>% 
    mutate(idx = row_number()) %>% 
    select(idx, everything()) # %>% 
    # mutate_at(vars(is_accepted), ~ifelse(. == 1, TRUE, FALSE))
  
  # if(!is.null(n_burnin)) {
  #   idx_slice <- (n_burnin + 1):nrow(mat_mcmc)
  #   res_mcmc <- res_mcmc %>% slice(idx_slice) 
  # }
  
  res_mcmc
}

do_mcmc_mh_q1 <- function(..., .n_mcmc = n_mcmc_q1, .n_burnin = n_burnin_q1) {
  .do_mcmc_mh_q1(n_mcmc = .n_mcmc, n_burnin = .n_burnin, ...)
}

```

```{r q1-a-3, include=T, echo=T, eval=T}
res_mcmc_mh_q1 <- do_mcmc_mh_q1()
res_mcmc_mh_q1
```

```{r q1-a-viz-debug-1, include=F, echo=F, eval=F}
# res_mcmc_mh_q1 %>% filter(is_accepted)
res_mcmc_mh_q1 %>% 
  ggplot() +
  aes(x = idx, y = theta_current) +
  geom_point() +
  geom_line()

res_mcmc_mh_q1 %>% 
  sample_frac(0.1) %>% 
  # select_if(is.numeric) %>% 
  gather(key = 'key', value = 'value', -idx) %>% 
  ggplot() +
  aes(x = idx, y = value) +
  geom_point() +
  facet_wrap(~key, scales = 'free')

res_mcmc_mh_q1 %>% 
  # select_if(is.numeric) %>% 
  gather(key = 'key', value = 'value', -idx) %>% 
  ggplot() +
  aes(x = value) +
  geom_histogram() +
  facet_wrap(~key, scales = 'free')
```

```{r q1-a-viz-1-prep, include=F, echo=F, eval=T}
path_viz_q1 <- 'viz_mcmc_mh_q1.png'
w_viz_q1 <- 8
h_viz_q1 <- 5
eval_viz_q1 <- !fs::file_exists(path_viz_q1)
```

```{r viz-funcs, include=T, echo=T, eval=T}
theme_custom <- function(...) {
  theme_light() +
    theme(
      axis.title.x = element_text(hjust = 1),
      axis.title.y = element_text(hjust = 1),
      ...
    )
}
```

```{r q1-a-viz-1, include=T, echo=T, eval=eval_viz_q1}
viz_mcmc_mh_q1 <-
  res_mcmc_mh_q1 %>% 
  ggplot() +
  aes(x = theta_prop) +
  geom_histogram() +
  scale_y_continuous(labels = scales::comma) +
  theme_custom() +
  labs(
    x = 'theta',
    y = 'Frequency'
  )
viz_mcmc_mh_q1
```

```{r q1-a-viz-1-export, include=F, echo=F, eval=eval_viz_q1}
teproj::export_ext_png(
  viz_mcmc_mh_q1,
  path = path_viz_q1,
  units = 'in',
  width = w_viz_q1,
  height = h_viz_q1
)
```

```{r q1-a-viz-1-show-1, include=T, echo=T, eval=F}
viz_mcmc_mh_q1
```

```{r q1-a-viz-1-show-2, include=T, echo=F, eval=T}
knitr::include_graphics(path_viz_q1)
```

# 2. Gibbs Sampler and High/Low Protein Diet in Rats.

## Instructions

## Response

### a


```{r q2-ab-1, include=T, echo=T, eval=T}
# Data, constants, hyperparameters, and functions.
y_1 <- c(134, 146, 104, 119, 124, 161, 107, 107, 83, 113, 129, 97, 123)
y_2 <- c(70, 118, 101, 85, 107, 132, 94)

n_mcmc_q2 <- 10000
n_burnin_q2 <- 500

theta_1_0 <- 110
theta_2_0 <- theta_1_0
tau_1_0 <- 1 / 100
tau_2_0 <- tau_1_0
a_1_0 <- 0.01
a_2_0 <- a_1_0
b_1_0 <- 4
b_2_0 <- b_1_0

.compute_mu_new <- function(mu_i, tau_i, mu_0, tau_0, y_sum, n_obs) {
  mu_tau_0 <- mu_0 * tau_0
  mu_rnorm_num <- tau_i * y_sum + mu_tau_0
  mu_rnorm_den <- tau_0 + n_obs * tau_i
  mu_rnorm <- mu_rnorm_num / mu_rnorm_den
  sigma2_rnorm <- 1 / (tau_0 + n_obs * tau_i)
  sigma_rnorm <- sqrt(sigma2_rnorm)
  rnorm(1, mu_rnorm, sigma_rnorm)
}

y_sum_1 <- sum(y_1)
y_sum_2 <- sum(y_2)
n_obs_1 <- length(y_1)
n_obs_2 <- length(y_2)
compute_mu_new_1 <-
  function(mu_i,
           tau_i,
           .mu_0 = theta_1_0,
           .tau_0 = tau_1_0,
           .y_sum = y_sum_1,
           .n_obs = n_obs_1) {
    .compute_mu_new(
      mu_i,
      tau_i,
      mu_0 = .mu_0,
      tau_0 = .tau_0,
      y_sum = .y_sum,
      n_obs = .n_obs
    )
  }

compute_mu_new_2 <-
  function(mu_i,
           tau_i,
           .mu_0 = theta_2_0,
           .tau_0 = tau_2_0,
           .y_sum = y_sum_2,
           .n_obs = n_obs_2) {
    .compute_mu_new(
      mu_i,
      tau_i,
      mu_0 = .mu_0,
      tau_0 = .tau_0,
      y_sum = .y_sum,
      n_obs = .n_obs
    )
  }

.compute_tau_new <- function(mu_new, y, a_0, b_0, n_obs) {
  shape_rgamma <- a_0 + 0.5 * n_obs
  rate_rgamma  <- b_0 + 0.5 * sum((y - mu_new) ^ 2)
  rgamma(1, shape = shape_rgamma, rate = rate_rgamma) 
}

compute_tau_new_1 <-
  function(mu_new,
           .y = y_1,
           .a_0 = a_1_0,
           .b_0 = b_1_0,
           .n_obs = n_obs_1) {
    .compute_tau_new(
      mu_new = mu_new,
      y = .y,
      a_0 = .a_0,
      b_0 = .b_0,
      n_obs = .n_obs
    )
  }

compute_tau_new_2 <-
  function(mu_new,
           .y = y_2,
           .a_0 = a_2_0,
           .b_0 = b_2_0,
           .n_obs = n_obs_2) {
    .compute_tau_new(
      mu_new,
      y = .y,
      a_0 = .a_0,
      b_0 = .b_0,
      n_obs = .n_obs
    )
  }

theta_1_init <- mean(y_1) # theta_1_0
theta_2_init <- mean(y_2) # theta_1_i
tau_1_init <- 1 / sd(y_1) # tau_1_0
tau_2_init <- 1 / sd(y_2) # tau_1_i
```

```{r q2-ab-2, include=T, echo=T, eval=T}
# Main function.
.do_mcmc_gibbs_q2 <-
  function(...,
           n_mcmc = 10000,
           n_burnin = 1000,
           theta_1_init = 0,
           theta_2_init = 0,
           tau_1_init = 1,
           tau_2_init = 1) {
    stopifnot(is_likeinteger(n_mcmc))
    if (!is.null(n_burnin)) {
      stopifnot(is_likeinteger(n_burnin))
      stopifnot(n_burnin < n_mcmc)
    }
    stopifnot(is.numeric(theta_1_init))
    stopifnot(is.numeric(theta_2_init))
    stopifnot(is.numeric(tau_1_init))
    stopifnot(is.numeric(tau_2_init))
    cols_mat_mcmc <- c('theta_1', 'theta_2', 'tau_1', 'tau_2')
    mat_mcmc <- matrix(nrow = n_mcmc, ncol = length(cols_mat_mcmc))
    
    colnames(mat_mcmc) <- cols_mat_mcmc
    theta_1_i <- theta_1_init
    theta_2_i <- theta_2_init
    tau_1_i <- tau_1_init
    tau_2_i <- tau_2_init
    y_1_sum <- sum(y_1)
    y_2_sum <- sum(y_2)
    n_1_obs <- length(y_1)
    n_2_obs <- length(y_2)
    for (i in 1:n_mcmc) {
      theta_1_new <-
        compute_mu_new_1(
          mu_i = theta_1_i,
          tau_i = tau_1_i
        )
      
      theta_2_new <-
        compute_mu_new_2(
          mu_i = theta_2_i,
          tau_i = tau_2_i
        )
      
      tau_1_new <-
        compute_tau_new_1(
          mu_new = theta_1_new
        )
      
      tau_2_new <-
        compute_tau_new_2(
          mu_new = theta_2_new
        )
      mat_mcmc[i,] <-
        c(theta_1_new, theta_2_new, tau_1_new, tau_2_new)
      
      theta_1_i <- theta_1_new
      tau_1_i <- tau_1_new
      theta_2_i <- theta_2_new
      tau_2_i <- tau_2_new
      
    }
    res_mcmc <-
      mat_mcmc %>% 
      as_tibble() %>% 
      mutate(idx = row_number()) %>% 
      select(idx, everything())
    
    if (!is.null(n_burnin)) {
      idx_slice <- (n_burnin + 1):nrow(mat_mcmc)
      res_mcmc <- res_mcmc %>% slice(idx_slice)
    }
    
    res_mcmc
  }

do_mcmc_gibbs_q2 <-
  function(...,
           .n_mcmc = n_mcmc_q2,
           .n_burnin = n_burnin_q2,
           .theta_1_init = theta_1_init,
           .theta_2_init = theta_2_init,
           .tau_1_init = tau_1_init,
           .tau_2_init = tau_2_init) {
    .do_mcmc_gibbs_q2(
      n_mcmc = .n_mcmc,
      n_burnin = .n_burnin,
      theta_1_iniit = .theta_1_init,
      theta_2_init = .theta_2_init,
      tau_1_init = .tau_1_init,
      tau_2_init = .tau_2_init,
      ...
    )
  }
# TODO: Remove these lines.
# undebug(compute_tau_new_1)
# res_mcmc_gibbs_q2 <- do_mcmc_gibbs_q2()
```

```{r q2-ab-3, include=T, echo=T, eval=T}
res_mcmc_gibbs_q2 <- do_mcmc_gibbs_q2()
res_mcmc_gibbs_q2
```

```{r q2-ab-4, include=F, echo=F, eval=F}
res_mcmc_gibbs_q2 %>% summarise_all(mean)
```

```{r q2-ab-viz-2-prep, include=F, echo=F, eval=T}
path_viz_q2 <- 'viz_mcmc_gibbs_q2.png'
w_viz_q2 <- w_viz_q1
h_viz_q2 <- h_viz_q1
eval_viz_q2 <- !fs::file_exists(path_viz_q2)
```

```{r q2-ab-viz-1, include=T, echo=T, eval=eval_viz_q2}
viz_mcmc_gibbs_q2 <-
  res_mcmc_gibbs_q2 %>% 
  gather(key = 'key', value = 'value', -idx) %>% 
  ggplot() +
  aes(x = value) +
  geom_histogram() +
  scale_y_continuous(labels = scales::comma) +
  theme_custom() +
  facet_wrap(~key, scales = 'free') +
  labs(
    x = NULL,
    y = 'Frequency'
  )
```

```{r q2-ab-viz-1-export, include=F, echo=F, eval=eval_viz_q2}
teproj::export_ext_png(
  viz_mcmc_gibbs_q2,
  path = path_viz_q2,
  units = 'in',
  width = w_viz_q2,
  height = h_viz_q2
)
```

```{r q2-ab-viz-1-show-1, include=T, echo=T, eval=F}
viz_mcmc_gibbs_q2
```

```{r q2-ab-viz-1-show-2, include=T, echo=F, eval=T}
knitr::include_graphics(path_viz_q2)
```

### b

```{r q2-b-1, include=T, echo=T, eval=T}
res_mcmc_gibbs_q2_calc <-
  res_mcmc_gibbs_q2 %>%
  mutate(
    theta_diff = theta_1 - theta_2
  ) %>% 
  mutate(
    h_0 = ifelse(theta_diff > 0, TRUE, FALSE)
  )

```

```{r q2-b-2, include=T, echo=T, eval=T}
h_0_q2 <-
  res_mcmc_gibbs_q2_calc %>% 
  summarise(h_0 = sum(h_0), n = n()) %>% 
  mutate(frac = h_0 / n)
h_0_q2
```

### c

```{r q2-c-0, include=F, echo=F, eval=F}
# compute_equi_credible_set <- function(mu, sigma, level = 0.95) {
#   alpha <- 1 - level
#   alpha_half <- alpha / 2
#   q_l <- alpha - alpha_half
#   q_u <- level + alpha_half
#   res <-
#     c(
#       l = qnorm(q_l, mean = mu, sd = sigma),
#       u = qnorm(q_u, mean = mu, sd = sigma)
#     )
# }
```

```{r q2-c-1, include=T, echo=T, eval=T}
equi_credible_set_q2 <-
  res_mcmc_gibbs_q2_calc %>% 
  summarise(
    mu = mean(theta_diff), 
    sd = sd(theta_diff)
  ) %>% 
  mutate(
    l = qnorm(0.025, mu, sd),
    u = qnorm(0.975, mu, sd)
  )
equi_credible_set_q2
```



